\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

We provide several data-driven control design frameworks for contact-rich
robotic systems.
%
These systems exhibit continuous state flows and discrete state transitions,
which are governed by distinct equations of motion.
%
Hence, it is difficult to design a single policy that can control the system in
all modes.
%
Typically, hybrid systems are controlled by multi-modal policies, each manually
triggered based on observed states.
%
However, as the number of potential contacts increase, the number of policies
can grow exponentially and the control-switching scheme becomes too complicated
to parameterize.
%
To address this issue, we design contact-aware data-driven controllers given by
deep-net mixture of experts.
%
This architecture learns switching-control scheme that can achieve the
desired overall performance of the system, and a gating network, which
determines the region of validity of each expert, based on the observed states
%


Additionally, we address the adverse effects of model uncertainties in the
control of contact-rich robots.
%
Lack of accurate environmental models can misrepresent the effects of contact
forces on the system.
%
Policies designed from such models can lead to poor performance or even
instability.
%
In particular, we demonstrate the effects of system parameter uncertainties and 
measurement errors on the overall performance of the system.
%
Then, we design data-driven stochastic controllers that combine the stability
properties of passivity-based control with the robustness properties of Bayesian
learning.
%


