\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

First and foremost, I want to thank my parents, Ayele and Mulu, for their hard
work in raising my sisters and myself in a nurturing and disciplined
environment.
%
I draw my inspiration from my mother, who instilled in me the love for education
and the desire to fulfill my potential. 
%
Thank you to all of my aunts and uncles, especially Bezu, Ayu, Hirut, Abraham,
and Legesse, who attended to my every need as a foreign student and contributed
towards my financial stability as I pursued my education in the US.  
%
My sisters, Rahel, Hermela, and Semawit are my greatest supporters and the
backbones of my mental well-being; thank you for tirelessly playing your
sisterly roles from thousands of miles away.
%
Thank you to all of my friends in Ethiopia, \textit{my sisters not by blood but by
bond}, from whom I find the security of life-long friendships preserved even in
long distances.
%
I am also thankful for the friends I made in the US, with whom I exchanged
culture and experienced true humanity in their love and affection. 



I am most grateful for my teachers, especially from middle and high school, for
giving me a strong foundation in my education, without which I would have
struggled tremendously.
%  
I extend a special thanks to my middle school mathematics teacher, Michael (Gash
Miki), who fostered discipline and rational thinking in me at such a young age.
%
I am indebted to all of my STEM instructors throughout my life, espcially Henok,
Tsegaye, Zimbelachew, Sintayehu, Dr Clare Fitzpatrick, Dr John Chiasson and Dr
Krishna Pakala for making me the engineer I am today.
%
This work would not have been possible without my advisor, Dr Aykut Satici,
whose endless enthusiams for all things STEM is inspiring.
%
Last, but definitely not least, thank you to my lab mates, past and present,
such as Dr Wankun Sirichotiyakul, Ryan Montrose, Master Alex Peterson,
Christopher Dagher (\textit{Engineer to the Stars}), Yafa Benavidez, Oliver
MacDonald and Chandika Silva; you make work and life so much fun.
